\documentclass{article}
\usepackage{alexconfig}
\title{Style Guide}

\begin{document}
\maketitle

\section{Philosophy}
This section of the style guide details the mindset I should have when contributing to this project.

The goal of this project is to catalog the subjects that I have learned, to assist learning by ensuring I grasp concepts through restating them in the study guides, and to provide clear, concise, and easy to read guides for myself and others to use.

There are four key tenets that this project abides by:

\begin{enumerate}
    \item Readability: Do not use any unnecesary jargon or confusing language. State and explain all concepts so that they are understandable even if you lobotomized half of your brain.
    \item Thoroughness: Proofs should be as thorough as they were in the book, or even more. Take the time to explain every step of thinking when solving problemss.
    \item Completeness: A completed study guide should cover every topic that the book does. Do not miss content.
    \item Originality: Make sure all content is stated in your own words, even if it is clunky and not as eloquent. Otherwise, you are copying from the book, and not actually learning.
\end{enumerate}

\section{Organization}
This section of the style guide details the top level organization and file structure of the document.

The "Study Guides" directory contains all of the study guides in this project. In addition, study guides get sorted (in increasing order of specificity), subjects, books, chapters, and lessons.

Study guides are usually templated around single lessons from books. The LaTeX file of a study guide should be placed in a lesson folder, and be named "notes.tex". The lesson folder should be named with a number corresponding to the lesson number of the study guide.

The lesson folders corresponding to a single chapter in a book should be placed in a chapter folder, which is named with a number corresponding to the chapter in the book.

Chapter folders should be placed in a book folder, which should be named according to the book's authors.

Book folders should be placed in 

\section{Formatting}
This section of the style guide details the look, feel, and structure within a given document.
\end{document}