\documentclass{article}
\usepackage{alexconfig}
\title{4.7: Poisson Random Variable}

\begin{document}
\maketitle

\begin{definition}[Poisson Random Variable]
The \textbf{poisson random variable} is a tool used to approximate binomial random variables with a large number of independent trials. It relies on a parameter $\lambda = np$ where $n$ is the number of independent trials and $p$ is the chance that each trial turns positive.  
\end{definition}

Lets derive the poisson random variable from the definition of a binomial random variable. Suppose $X$ is a binomial random variable with $n$ trials and $p$ chance for each success. Then, $$P\{X = i\} = \binom{n}{i}p^i(1-p)^{n-i}$$Expanding the binomial, we get $$P\{X = i\} = \frac{n!}{i!(n-i)!}p^i (1-p)^{n-i}$$If we let $\lambda = np$ and substitute in, we get $$P\{X = i\} = \frac{n!}{i!(n-i)!}(\frac{\lambda}{n})^i (1-\frac{\lambda}{n})^{n-i}$$And if we cancel $n!$ with $(n-i)!$ in our binomial and split the last term, we get $$P\{X = i\} = \frac{n(n-1)(n-2)...(n-i+1)}{i!}\frac{\lambda^i}{n^i}\frac{(1-\frac{\lambda}{n})^n}{1-\frac{\lambda}{n}^i}$$Rearranging, we get $$P\{X = i\} =\frac{n(n-1)(n-2)...(n-i+1)}{n^i}\frac{\lambda^i}{i!}\frac{(1-\frac{\lambda}{n})^n}{1-\frac{\lambda}{n}^i}$$Notice that when $n$ is large and $\lambda$ is moderate, $$(1-\frac{\lambda}{n})^n \cong e^{-\lambda}$$ $$\frac{n(n-1)(n-2)...(n-i+1)}{n^i} \cong 1$$ $$(1-\frac{\lambda}{n})^i \cong 1$$This means that under these conditions, $$P\{X = i\} \cong e^{-\lambda}\frac{\lambda^i}{i!}$$

Some use cases of a poisson random variable include:

\begin{enumerate}
    \item Determining the amount of misprints in a set of pages in a book.
    \item The number of people who survive to age 100.
    \item The amount of people whose birthdays match a given day.
\end{enumerate}


\begin{example}
Suppose the probability an item produced by a machine is defective is $p = .1$. Out of a sample of $10$ items, what is the probability it will contain at most $1$ defective item?
\end{example}

\begin{solution}
Let $X$ be a poisson random variable with $\lambda = np = .1 * 10 = 1$ of a sample having $x$ defective items. We want to find $P\{X = 0\} + P\{X = 1\}$ Plugging in, we get that $$P\{X = 0\} = e^{-1}\frac{\lambda^0}{0!} = \frac{1}{e}$$$$P\{X = 1\} = e^{-1}\frac{\lambda^1}{1!} = \frac{1}{e}$$ So in total, the chance of having 1 defective part or less is $$\frac{2}{e}$$
\end{solution}

Since the poisson random variable is an approximation of the binomial random variable, the expected value is $$E[X] = np = \lambda$$ and $$\text{Var}[X] = np = \lambda$$
\end{document}